---
title: "MATH 185 - Computational Statistics - Homework 1"
author: "Dukki Hong"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_notebook: default
geometry: margin=2cm
Section: A04
PID: A98058412
---

```{r, include=FALSE}
#install.packages('tinytex',repos = "http://cran.us.r-project.org")
#options(tinytex.verbose = TRUE)
```

## Problem 1


### Part A


```{r}
chisq.power <- function(k,t,n,B = 2000){
  R <- vector(mode='numeric', length = B)     #create an empty binary vector of length B
  
  null <- numeric(2*k)                        #data under the null, uniformly distributed
  
  for (i in 1:k){                             #Generate data for Observed counts
    null[i] <- 1/(2*k) + t
  }
  for (i in (k+1):(2*k)){
    null[i] <- 1/(2*k) - t
  }
  
  #Monte Carlo Simulation
  for(i in 1:B){
    sim = sample(1:(2*k), n, replace = T, prob = null) 
    sim = table(sim)
    if (chisq.test(sim)$p.value <= 0.05){
      R[i] = 1
    }
    else{
      R[i] = 0
    }
  }
  
  return(sum(R)/B) #return proportion of correctly rejecting H_0 over H1
}
```


### Part B
```{r, warning=FALSE}
k = 6
t = seq(0+0.005, 1/(2*k) - 0.005, by =0.005) #range of values for t
n = c(100,300,500,1200) #set different values for n

pwr = numeric(length(t)) #powers

ylab = ('Power')  
main = ('Power Curve')
col = ('red')

par(mfrow = c(2,2))
for (j in 1:length(n)) { #for loop to calculate over all n values
  for (i in 1:length(t)) {
    pwr[i] <- chisq.power(k, t[i], n[j])
    i = i + 1
    }
  plot(t, pwr, type = 'l', xlab = 't', ylab = ylab, main = main, col=col)
}
```



## Problem 2

Are the chance of a baby being born a girl the same across counties in California? Go
to http://wonder.cdc.gov/natality.html. In Section 1, choose Gender and County. In Section
2, choose the state to be California. In Section 4, choose the year 2017. Then click on Send"
anywhere. Enter the data in R. You can do so by hand or click on Export", edit the resulting text
le and then use the function read.table. (You will have to edit the TXT le in order to read it into
R.) Save the dataset as an RDA le named natality-california-2017.rda. Do this as a preprocessing.
Start your code by loading the dataset using the function load.

Use these data to answer the question the best you can. Start by formalizing the question into a hypothesis testing problem, display relevant summary statistics and graphics, and then perform an appropriate test. Conclude with a sentence or two.

```{r, warning=FALSE}
mydata <- read.table("~/Desktop/UCSD MATH/Math.185/Natality, 2007-2017.txt", header = TRUE)
save(mydata, file="~/Desktop/UCSD MATH/Math.185/natality-california-2017.rda")

load("~/Desktop/UCSD MATH/Math.185/natality-california-2017.rda") 

mydata = data.frame(mydata) #change mydata into a dataframe
mydata = subset(mydata, select=c('Gender', 'County.Code','Births')) #remove unnecessary columns

#Are the chance of a baby being born a girl the same across counties in California?
#Parameter of interest: chance of baby being born a girl (p_i)
#H_0: chance of baby being born a girl SAME across counties in CA
#H_1: chance of baby being born a girl DIFFERENT across counties in CA

mydata = mydata[order(mydata$County.Code),] #sort out 'County.Code' column in an increasing order

female = mydata[seq(1,72,2),] #sort out mydata with females only
male = mydata[seq(2,72,2),] #sort out mydata with males only

f = matrix(female$Births, nrow = 1, ncol=36) #turn data into matrices
m = matrix(male$Births, nrow = 1, ncol=36)

X = rbind(f,m) #combine the two matrices into one matrix
row.names(X) = c('Female','Male') #Label row names 
  
chisq.test(X)

cat('value of X-Squared is', '41.285,', 'df is 35', 'and p-value is:', '0.215.')
```


*Now we proceed to showing a graphical representation of the data*
```{r}
col = c('pink','skyblue') #set color
ylab = ('Number of births') #Y-label name
barplot(X, legend=T, col=col, ylab=ylab, ylim=c(0,120000)) 
barplot(X, , beside= T, legend=T, col=col, ylab=ylab)
```
<font color=red>We see similarities between Male and Female groups in the graph above</font>

## Problem 3

Write a function *chisq.perm.test*(tab;B = 2000) implementing the chi-squared test of independence calibrated by permutation. The inputs are a table of joint counts (without totals) and number of permutations to be done. Have the test return the p-value. Test your function on the *HairEyeColor* dataset.

```{r}
chisq.perm.test <- function(tab, B=2000){
  obs = chisq.test(tab)$stat #observed statistic from the Chisqd test
  
  nrow = nrow(tab) #Number of rows
  ncol = ncol(tab) #Number of columns
  
  rows = numeric(nrow) #create a 0 vector with n number of rows
  totrow = numeric(0) #create a vector with 0's
  
  for (i in 1:nrow){ #for-loop to fill up the vectors defined above
    rows[i] = sum(tab[i,]) #summing up each of the rows
    totrow = c(totrow, rep.int(i, times=rows[i]))
  }
  
  #do the same with the columns
  cols = numeric(ncol) #empty vector with n columns
  totcol = numeric(0) #0 vector
  
  for (i in 1:ncol){
    cols[i] = sum(tab[,i]) #summing up each of the columns
    totcol = c(totcol, rep.int(i, times=cols[i]))
  }
  
  D = numeric(0) #vector for permutations for the Chisqd test statistic
  tot = sum(tab[,]) #total number of counts
  
  #####################################
  
  for (b in 1:B){
    x = totrow
    y = sample(totcol, tot, replace=FALSE) #sample column totals 'tot' times without replacement
    matperm = cbind(x,y) #combine columns of x and y
    
    matpermcount = matrix(0, nrow=nrow, ncol=ncol) #create a 0 matrix with n number of rows and columns
    
    for(x in 1:nrow){ #rows from 1 - 4
      for(y in 1:ncol){ #rows from 1 - 4
        for(m in 1:tot){ #loop for going through all the rows in matperm
          if(all(matperm[m,] == c(x,y)) == TRUE){ #ordered pairs such as (1,1), (2,2) and etc...
            matpermcount[x,y] = matpermcount[x,y] + 1
          }
        }
      }
    }
  
    D[b] = chisq.test(matpermcount)$stat #chisq.test statistic for a given permutation
  
    #Check if chisq.test stat >= observed stat
    count = 0 #counts to check (chisq.test stat > observed)
    if (D[b] >= obs){
      count = count + 1
    }
    return((count+1)/(B+1))
  }
}
```
*Now we check if the function works*
```{r}
tab = apply(HairEyeColor, c(1,2), sum)
tab
chisq.perm.test(tab)
cat('We get that the Chi-squared test statistic is equal to :',chisq.perm.test(tab))
```



## Problem 4

Go to the following webpage: https://catalog.data.gov/dataset/school-improvement-2010-grants 
Download the dataset. The description is on the webpage. To read it into R, use the function
read.csv. Remove the schools from Rhode Island as their selected models are missing. Do this on your own to practice reading datasets into R. Save the dataset as an RDA file named school-improvement-2010.rda. Do this as a preprocessing. Start your code by loading the dataset using the function load.

```{r}
#Loading and cleaning data
data = read.csv('~/Desktop/UCSD MATH/Math.185/userssharedsdfschoolimprovement2010grants.csv', head=T)

head(data)
#data[sample(nrow(data),5),] 5 random sample of the data
data[which(data$State=='RI'),] #we see that schools in RI have missing models
data = data[which(data$State != 'RI'),] #remove all the schools from RI (Rhode Island)

save(data, file="~/Desktop/UCSD MATH/Math.185/school-improvement-2010.rda") #save dataset as an rda file
```

## Problem 4
We are interested in the following question: Is there an association between the model that each school selected and the state where the school was located at that time?

### Part A

Explore this question with one or several appropriate plots. Then formulate the question into a hypothesis testing problem and perform a test. Conclude with some brief comments.

#<font color=red>H_0: There is no association between the model that each school selected and the state where the school was located at that time, i.e. independence.</font>

#<font color=red>H_1: There is association between the model that each school selected and the state where the school was located at that time, i.e. dependence.</font>

<font color=red>Since significance level was not mentioned, we proceed and set a =0.05.</font>
```{r, warning=FALSE}
load("~/Desktop/UCSD MATH/Math.185/school-improvement-2010.rda")
data = data[,c('State','Model.Selected')] #sort out the state and Model.selected columns
table = table(data$State, data$Model.Selected) #create a table using the two columns
table = table[c(1:38,40:50), c(2:5)]

xlab = ('Models')

barplot(table, beside=T, xlab=xlab)
barplot(table, xlab=xlab)

chisq.test(table)
cat('p-value is',chisq.test(table)$p.value, '< 0.05=a', 'so we reject the null hypothesis. This means that there is association between the model that each school selected and the state where the school was located at that time. Hence the States and Models are dependent.')
```


### Part B

In this particular case, is the method of Problem 3 applicable? If so, apply it and compare with the previous test that you performed.

<font color=red>It is applicable to use the method from Problem 3 in this case.</font>
```{r, warning=FALSE}
chisq.perm.test(table)
cat('Despite the fact that the p-value we obtain is:', chisq.perm.test(table), "which is larger than the p-value from the Chi-Squared test, but not large enough to keep H_0. Hence we still reject the null hypothesis and conclude that there is no association between the state and the model.")

fisher.test(table, simulate.p.value=TRUE)
cat("Using 'fisher.test' function, we also get a p-value similar to that of the 'chisq.test':", fisher.test(table, simulate.p.value=TRUE, B=2000)$p.value)
```
<font color=red>Despite the fact that the p-value from 'chisq.perm.test' is larger than the p-value from the Chi-Squared test but it is not too large to keep the null hypothesis. Hence we still have the fact that there is no association between the state and the model.</font>

